Aligning species range data to better serve science and conservation

### 2016-01-23 manuscript-main.Rmd
updating the manuscript...

---
title: "Aligning species range data to better serve science and conservation"
author: "Casey O'Hara, Jamie Afflerbach, Benjamin S. Halpern"
output:
  html_document:
    highlight: haddock
    number_sections: yes
    theme: spacelab
    toc: no
  pdf_document:
    number_sections: yes
    toc: no
  word_document: default
---

>>>>

---
title: "Aligning species range data to better serve science and conservation"
author: "Casey O'Hara, Jamie Afflerbach, Benjamin S. Halpern"
output:
  pdf_document:
    number_sections: yes
    toc: no
  html_document:
    highlight: haddock
    number_sections: yes
    theme: spacelab
    toc: no
  word_document: default
---

<!--

from PNAS (http://www.pnas.org/site/authors/preparation.xhtml):

PNAS generally uses a two-column format averaging 67 characters, including spaces, per line. The maximum length of a regular research article is 6 pages (or 10 pages for PNAS Plus articles), including all text, spaces, and the number of characters displaced by figures, tables, and equations.

An online submission tool provides authors with an estimation of whether their manuscript fits within the PNAS length requirements (see Length Estimate FAQ). Use the following guidelines to determine whether your manuscript exceeds the page-length requirements before submission:

Text: When submitting tables, figures, and/or equations in addition to text, keep the text for your manuscript under 39,000 characters (including spaces).
Figures: Calculated at 180 characters per cm in height for one column and 360 characters per cm in height for two columns; a figure wider than 8.7 cm is two columns wide.
Tables and equations: Calculated at 60 characters per line for one column and 120 per line for two columns; a table with more than 60 characters per line is two columns wide.
If the text in a figure or table is deemed illegible for production, the figure or table may be enlarged to improve legibility. Add the word count to the character count to obtain a character count that includes spaces.

See also: http://www.pnascentral.org/html/pnas_length_estimate_instructions.html

-->

### 2016-01-23 ms_2_intro_res_disc.Rmd
updated figures without titles; boxplot = violin plot
updating the manuscript...

To understand the implications of differences between the AquaMaps and IUCN datasets, we compared how each data source represents the global spatial and taxonomic distribution of species.  Most notably, AquaMaps includes range maps for many more species (currently `r format(n_spp_am, big.mark = ',')`; `r p_spp_am`% of total), such that most global analyses related to biodiversity to date have used AquaMaps (IUCN range map data exist for only `r format(n_spp_iucn, big.mark = ',')` species).  For the `r format(n_spp_both, big.mark = ',')` (`r p_both_all`% of total) species mapped in both datasets, we examined how well the maps align and determined several issues that lead to misalignment between predicted species distributions, and outline possible improvements.

@@

-----
<!--
## Overview of AquaMaps and IUCN datasets

@@

The IUCN publishes species range maps as polygon shapefiles, bundled by taxonomic groups (REF).  Using a geographic information system (GIS), experts outline spatial polygons to represent a given species' "limits of distribution" (IUCN 2014) - essentially a refined extent of occurrence, based on observation records and refined by expert understanding of the species' range and habitat preferences.  AquaMaps models species distribution maps based environmental preferences (e.g. temperature, depth, salinity) deduced from occurrence records, published species databases such as FishBase, and expert knowledge.  The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a 0.5 degree grid, creating a global raster of probability of occurrence for each species.

The two datasets have notably different taxonomic (Fig. 1A) and regional (Fig. 1B,C) coverage. IUCN-mapped species focus on tropical latitudes and away from the Atlantic and Eastern Pacific compared to AquaMaps-mapped species. These differences can be mitigated by using both datasets, but the underlying methodological differences complicate such direct comparisons, and furthermore ___requires resolving differences (or choosing one source over the other) for the `r format(n_spp_both, big.mark = ',')` shared species.___

@@

___BSH suggestion___ To explore differences in species distribution and range between the two datasets, we plotted the distribution alignment (where on the map) against the ratio of extent (how much of the map) for each shared species (Fig. 2a). __Overall we see a negative linear relationship, where increasing similarity in extent (extent ratio approaches 100%) correlates with a decrease in alignment.  [explain/discuss].__

@@

The extent-mismatched maps contained in the upper left quadrant (n = `r n_q2`) provide a richer set of examples to understand the fundamental differences between the datasets.  For many of these species, while the spatial distribution is similar, the IUCN range predicts a significant buffer zone beyond the AquaMaps range.  Corals dominate this quadrant (`r n_q2_corals`; `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in this quadrant).  Most corals and reef-associated organisms prefer shallower waters, and we suspect that many of these disagreements in extent can be explained by bathymetry: seafloor depth is explicitly included in AquaMaps models, while depth is frequently ignored as a factor in IUCN range maps, leading to overprediction of species range into areas too deep to support the species. (For examples and details, see SOM)

@@

The global Ocean Health Index (OHI) (Halpern et al. 2012, 2015), a composite index made up of 10 goals, uses species spatial distribution data and IUCN Red List conservation status to calculate biodiversity status for each of 221 exclusive economic zones. To maximize the number of represented species, spatial distributions were gleaned from both IUCN and AquaMaps datasets, preferring IUCN data for species in both sources.  OHI uses a probability threshold of 40% to determine species presence for AquaMaps data.  Species with no spatial data in either dataset were excluded, as were species with insufficient information to determine conservation status (including species listed as not evaluated or data deficient).

_Briefly summarize results_

Since the Ocean Health Index Species subgoal relies on spatial data from both datasets, the impacts of these threshold and preference changes will be somewhat muted (<a href = '#fig3'>Figure 3</a>).  When IUCN range maps are the preferred data source, only the subset of AquaMaps-only species will be affected by threshold changes; and when AquaMaps is the preferred source, the IUCN-only species will dampen the effect of a threshold change.

@@

working on the analysis...

* rerun based on old AquaMaps at 50% (to verify methodology gets a close result to paper), then new AquaMaps for updated baseline
    * consider also: filter on STATUS == 'Designated', and filter out those pesky fishery management areas and manatee speed zones...
* run with threshold at 0% (not all the same as original paper)
* run with filtering out data-poor species?
* run with IUCN maps instead of AquaMaps
* Which aspects of the analysis should we focus on?:
    * overall protection, for 4th subset of protected areas? or overall for all subsets? (see paper, table 1)
    * overall vs broken down by taxonomic groups? (see paper, figure 1)
    * spatial distribution of gap species (see paper, figure 2)

@@

_While the differences between species distributions described by AquaMaps and IUCN can drive different conclusions for conservation research, some of the mechanistic causes can be addressed._ ___oh god this section is so passive voice... arrrrgghhhh___

__For data providers:__

* The training manual for experts creating IUCN Red List maps for marine species suggests that "bathymetry can be used to delineate species' ranges limited by depth in the same way as elevation is used for terrestrial species" (REF https://www.conservationtraining.org/mod/scorm/player.php?a=1412&currentorg=M5L1_ORG&scoid=6492).  Unfortunately, this recommendation appears to be applied haphazardly.  Especially for demersal communities dependent on photosynthesis, bathymetry should be a primary consideration; even a low resolution cutoff of 200 meters would improve the utility of species range maps.

* The AquaMaps model is a powerful tool for quickly estimating species presence based on limited information; however, the model's output is only as good as the input data, which for marine species is often sorely limited.  AquaMaps encourages experts to review and comment on the predicted species distribution maps; lessons from the small sample of reviewed maps can be generalized, for example to other species in the same taxonomic group, to further refine the predictions, especially for data-poor species.

* In addition to environmental preferences and conditions, AquaMaps mapping parameters include area restrictions to help limit over-extrapolation of AquaMaps models.  Currently the area restrictions are denoted by FAO Major Fishing Areas (REF find a reference in AquaMaps somewhere); however, allowing area restrictions based on biogeographical criteria, a la MEOW, would likely have better resolution and predictive power, especially for data-poor species.

__For data users:__

>>>>

To understand the implications of differences between the AquaMaps and IUCN datasets, we compared how each data source represents the global spatial and taxonomic distribution of species.  Most notably, AquaMaps includes range maps for many more species (currently `r format(n_spp_am, big.mark = ',')` species; `r p_spp_am`% of total), such that most global analyses related to biodiversity to date have used AquaMaps (IUCN range map data exist for only `r format(n_spp_iucn, big.mark = ',')` species).  For the `r format(n_spp_both, big.mark = ',')` species (`r p_both_all`% of total) mapped in both datasets, we examined how well the maps align and determined several issues that lead to misalignment between predicted species distributions, and outline possible improvements.

@@

<!--
## Overview of AquaMaps and IUCN datasets

@@

The IUCN publishes species range maps as polygon shapefiles, bundled by taxonomic groups (REF).  Using a geographic information system (GIS), experts outline spatial polygons to represent a given species' "limits of distribution" (REF IUCN 2014) - essentially a refined extent of occurrence, based on observation records and refined by expert understanding of the species' range and habitat preferences.  AquaMaps models species distribution based environmental preferences (e.g. temperature, depth, salinity) deduced from occurrence records, published species databases such as FishBase, and expert knowledge.  The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a 0.5 degree grid, creating a global raster of probability of occurrence for each species (AquaMaps 2013)

The methodologies behind these datasets imply differences in prediction of species distribution due to errors of commission (falsely indicating species presence) and omission (falsely indicating species absence).  Geographic range data such as IUCN range maps generally include large commission errors, while predicted distribution models such as AquaMaps will likely include fewer commission errors but more omission errors.  Each type of error bears different implications for conservation goals (Rondinini, 2006).

<!-- Geographic ranges generate large commission errors because they assume homogenous species distributions. Predicted distribution data make explicit inferences on species occurrence and their commission and omission errors depend on model structure, on the omission of variables that determine species distribution and on data resolution. Omission errors lead to identifying networks of areas for conservation action that are smaller than required and centred on known species occurrences, thus affecting the comprehensiveness, representativeness and efficiency of selected areas. Commission errors lead to selecting areas not relevant to conservation, thus affecting the representativeness and adequacy of reserve networks. -->

The two datasets have notably different taxonomic (Fig. 1A) and regional (Fig. 1B,C) coverage. IUCN-mapped species focus on tropical latitudes and away from the Atlantic and Eastern Pacific compared to AquaMaps-mapped species. These differences can be mitigated by using both datasets, but the underlying methodological differences complicate such direct comparisons. ___and furthermore requires resolving differences (or choosing one source over the other) for the `r format(n_spp_both, big.mark = ',')` shared species.___

@@

To explore differences in species distribution and range between the two datasets, we plotted the distribution alignment (where on the map) against the ratio of extent (how much of the map) for each shared species (Fig. 2A).

@@

The extent-mismatched maps contained in the upper left quadrant (n = `r n_q2`) provide a richer set of examples to understand the fundamental differences between the datasets.  For many of these species, while the spatial distribution is similar, the IUCN range predicts a significant buffer zone beyond the AquaMaps range.  Corals dominate this quadrant (n = `r n_q2_corals`; `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in this quadrant).  Most corals and reef-associated organisms prefer shallower waters, and we suspect that many of these disagreements in extent can be explained by bathymetry: seafloor depth is explicitly included in AquaMaps models, while depth is frequently ignored as a factor in IUCN range maps, leading to overprediction of species range into areas too deep to support the species. (For examples and details, see SOM)

@@

The global Ocean Health Index (OHI) (Halpern et al. 2012, 2015), a composite index made up of 10 sustainable benefits provided by a healthy ocean, uses species spatial distribution data and IUCN Red List conservation status to calculate biodiversity status (scored from zero to 100) for each of the world's 221 exclusive economic zones. To maximize the number of represented species, OHI gleans spatial distributions from both IUCN and AquaMaps datasets, prioritizing IUCN data for the subset of species in both sources.  OHI uses a probability threshold of 40% to determine species presence for AquaMaps data.

We recalculated the OHI species status score under several scenarios to observe the impact of toggling the prioritized data set from IUCN to AquaMaps, and toggling the AquaMaps presence threshold from 40% to 0% (<a href = '#fig3'>Fig. 3</a>).  Most country scores varied by fewer than five points across the three scenarios; a few country scores decreased by as many as 12 points, indicating a more dire assessment of species conservation status.

_Since the Ocean Health Index Species subgoal relies on spatial data from both datasets, the impacts of these threshold and preference changes will be somewhat muted.  When IUCN range maps are the preferred data source, only the subset of AquaMaps-only species will be affected by threshold changes; and when AquaMaps is the preferred source, the IUCN-only species will dampen the effect of a threshold change._

@@

___working on the analysis...___

@@

Understanding the causes and implications of the differences between these two important datasets can help us better align the predictions of species distributions, increasing their utility for conservation goals.

__Suggestions for data providers:__

* The training manual for experts creating IUCN Red List maps for marine species suggests that "bathymetry can be used to delineate species' ranges limited by depth in the same way as elevation is used for terrestrial species" (REF https://www.conservationtraining.org/mod/scorm/player.php?a=1412&currentorg=M5L1_ORG&scoid=6492).  Unfortunately, this recommendation appears to be applied haphazardly.  Especially for demersal communities dependent on photosynthesis, bathymetry should be a primary consideration; even a simple cutoff at the photosynthetic limit of 200 meters would improve the utility of species range maps.
* The AquaMaps model is a powerful tool for quickly estimating species distribution based on limited information; however, the model's output is only as good as the input data, which for marine species is often sorely limited.  AquaMaps encourages experts to review and comment on the predicted species distribution maps; lessons from the small sample of reviewed maps can be generalized, for example to other species in the same taxonomic group, to further refine the predictions, especially for data-poor species.
* In addition to environmental preferences and conditions, AquaMaps mapping parameters include area restrictions to help limit over-extrapolation of AquaMaps models.  Currently the area restrictions are denoted by FAO Major Fishing Areas (REF find a reference in AquaMaps somewhere); however, allowing area restrictions based on biogeographical criteria, a la Marine Ecoregions of the World (), would likely have better resolution and predictive power, especially for data-poor species.

__Suggestions for data users:__

### 2016-01-23 ms_3_methods.Rmd
updating the manuscript...

### taxonomic distribution comparison

To examine the overall taxonomic distribution across the spatial datasets, we grouped species by taxonomic class and data source (IUCN, AquaMaps, or both), and examined the proportion of each class represented in each data source category.  We then filtered the species list to those that have been evaluated for the IUCN Red List of Threatened Species (in SOM).

___Currently this is also included in body - take out and leave in here?  Per BH: "Either cut or move to data description section of methods above."___: The IUCN releases spatial data sets when a taxonomic group (typically on the scale of order or family) has been comprehensively assessed, to guard against sample bias (though non-comprehensive datasets are available for reptiles and marine fish).  As such, spatial data for many taxonomic classes remain unavailable, and within a class, the assessed sub-groups may not represent the entire class.

### global spatial distribution comparison

To compare the spatial representation of the two datasets directly, we first rasterized the IUCN species polygons to the same 0.5° grid as the AquaMaps species maps; species presence within a grid cell was determined by any non-zero overlap of a species polygon with the cell, and species richness per cell was simply the count of the species present.  For the AquaMaps dataset, we determined per-cell species richness by counting all species with non-zero probability of occurrence, to best approximate the "extent of occurrence" generally indicated by IUCN maps.  We represented relative distribution of species richness for each dataset by plotting average species count against latitude and longitude.

_map pairs comparison_

Using genus and species binomials to identify paired maps, we selected the subset of marine species that have range maps in both IUCN and AquaMaps current native distribution (n = `r nrow(spp_map_pairs)`).  We used the same criteria as outlined above to determine species presence within a cell.

___from results/discussion: paraphrase or expand upon this___ Overlaying paired distribution maps for a given species, we calculated two dimensions of spatial alignment: _distribution alignment_, which we defined as the proportion of the smaller range intersecting the larger range; and _area alignment_, which we defined as the ratio of the smaller range to the larger range.

@@

@@

@@

>>>>

### comparison of taxonomic and regional distribution

To examine the overall taxonomic distribution across the spatial datasets (Fig. 1a), we grouped species by taxonomic class and data source (IUCN, AquaMaps, or both), and examined the proportion of each class represented in each data source category.

To compare the spatial representation of the two datasets directly, we rasterized the IUCN species polygons to the same 0.5° grid as the AquaMaps species maps; species presence within a grid cell was determined by any non-zero overlap of a species polygon with the cell, and species richness per cell was simply the count of the species present.  For the AquaMaps dataset, we determined per-cell species richness by counting all species with non-zero probability of occurrence, to best approximate the "extent of occurrence" generally indicated by IUCN maps.  We represented relative distribution of species richness for each dataset by plotting average species count against latitude and longitude (Fig. 1b, 1c).

### comparison of paired maps

Using genus and species binomials as a matching key, we selected the subset of marine species that have range maps in both IUCN and AquaMaps current native distribution.  To avoid double-counting, we removed subpopulations and species aliases.  We determined species presence within each spatial cell using the same criteria as outlined above.

Overlaying paired distribution maps for a given species, we defined and calculated _distribution alignment_ and _extent alignment_ and plotted these in Fig. 2:

@@

We visually inspected a random selection of paired distribution maps from each quadrant to identify possible mechanistic causes of misalignment.  To verify that IUCN predicted unsuitable habitat for depth limited species, we used QGIS (REF) to overlay a selection of IUCN and AquaMaps maps with a 200 meter bathymetry contour.

@@

Using methods and supplemental materials from OHI (Halpern et al. 2012, 2015), we modified the original code for OHI 2015 Species status (SPP) (REF), allowing for flexibility in prioritized data source and AquaMaps presence threshold.  We ran the SPP code three times, prioritizing IUCN over AquaMaps for a 0% threshold, and prioritizing AquaMaps over IUCN for both a 40% and 0% threshold.  We compared each of these to the output of the published OHI 2015 SPP model (prioritizing IUCN over AquaMaps at a 40% threshold).

@@

Based upon the methods described in Klein et al. 2015, we reconstructed the analysis using the subset of protected areas spatially covering a marine area and classified as IUCN I-IV ___(Klein paper, table 1, subset 4 - seems to be the one they focus on the most)___, converting the World Database of Protected Areas (WDPA, REF) dataset to a raster of 0.5° cells indicating proportion of protected area within the cell ___(consider also: filter on STATUS == 'Designated', and filter out those pesky fishery management areas and manatee speed zones...)___.  To verify our reconstruction we ran the analysis using the AquaMaps dataset available in April 2015 (REF) at a presence threshold of 50%.  We then updated the analysis using the most recent AquaMaps dataset (REF?) at 50% threshold

* Possible analyses:
    * run AquaMaps (new data) with threshold at 0%
        * already done in original paper, but redo it to compare to AM + IUCN scenario below
    * run with AquaMaps only, but filtering out data-poor species?
    * run with IUCN maps only in place of AquaMaps?
    * run with AquaMaps and IUCN both, using 0% threshold; which preferred? 
        * Should we clip corals (and coral-associated taxa) to 200 meter bathymetry, just for fun?
* Which aspects of the analysis should we focus on?:
    * overall vs broken down by taxonomic groups? (see paper, figure 1)
    * spatial distribution of gap species (see paper, figure 2)

### 2016-01-23 ms_4_figs.Rmd
updating the manuscript...

@@

### Figure 1

@@

>  __Figure 1(a):__ Number and proportion of species, listed by taxa, included in each dataset: IUCN, AquaMaps, or both.  AquaMaps encompasses a broader range of taxa than IUCN, while IUCN focuses on comprehensively assessing select taxonomic groups, typically at the level of order or family. Overlapping species are dominated by bony fishes (`r n_bonyfish` species, primarily tropical taxa) and corals (`r n_corals` species).

> __Figure 1 (b, c):__ Global marine species richness according to (b) AquaMaps dataset and (c) IUCN dataset.  The frequency plot to the right of each map shows relative species count per cell at each latitude; while both datasets peak in tropical latitudes near the equator, the frequency for IUCN maps drops quickly beyond 30°N and 30°S, while the frequency for AquaMaps remains robust well into temperate latitudes.  The frequency plot above each map shows relative species count at each longitude, showing a slight bias in the IUCN dataset away from the Atlantic and eastern Pacific compared to AquaMaps.

@@

![](figures/AreaRatiovsPercOverlap.png)

> __Figure 2 (a):__ For each paired-map species, we calculated two dimensions of spatial alignment: _distribution alignment_, which we defined as the proportion of the smaller predicted range intersecting the larger predicted range (where on the map); and _extent alignment_, which we defined as the ratio of the smaller range area to the larger range area (how much of the map).  For a species whose distribution is well understood and described in both datasets, we would expect to see a value near 100% for each dimension of alignment.

@@

> The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the extent of described ranges
(n = `r q1_count`; `r round(q1_count/nrow(spp_map_pairs) * 100, 1)` %).
> The upper left quadrant (quadrant 2) comprises species whose maps agree well in distribution, but disagree in extent
(n = `r q2_count`; `r round(q2_count/nrow(spp_map_pairs) * 100, 1)` %).
> The lower right quadrant (quadrant 3) includes species for which the paired maps generally agree in range extent, but disagree on where those ranges occur
(n = `r q3_count`; `r round(q3_count/nrow(spp_map_pairs) * 100, 1)` %).
> The lower left quadrant (quadrant 4) indicates species for which the map pairs agree poorly in both area and distribution
(n = `r q4_count`; `r round(q4_count/nrow(spp_map_pairs) * 100, 1)` %).

![](figures/barchart_spp_gp_quads.png)

> __Figure 2 (b):__ Spatial alignment of paired-map species by taxonomic group. ___REDO with all corals grouped together!___  _Also - is it distracting to use order names for some groups and not others?  Some of the groups had pretty long names ("acanthuridae" is way shorter than "surgeonfish, tangs, and unicornfish")_

<!--AquaMaps accounts for seafloor depth as a limiting factor, highly constraining the predicted habitat for photosynthesizing corals, while IUCN maps for coral species often include wide buffers that would extend far beyond appropriate depth for coral habitat, thus likely overestimating the actual species range (see SOM).  A similar mechanism may also drive poor extent alignment of other demersal and reef-associated organisms.-->

@@

> __Figure 3:__ Change in status score for the Species Subgoal within the global Ocean Health Index under three different scenarios.

@@

> * Scenario 1 shows the effect of reducing the presence threshold for AquaMaps presence.  Reducing the threshold will always increase the apparent range of a species, therefore the slight decrease in average score suggests increased spatial representation of threatened species.
> * Scenario 2 shows the effect of preferring AquaMaps data over IUCN, while maintaining the same presence threshold.  This will have different effects depending on the species; in general, AquaMaps ranges are smaller than IUCN ranges, so many but not all overlapping species will see a decrease in represented range.  The slight bump in mean score may indicate a small increase in spatial representation of low-risk species, a small decrease in spatial representation of high-risk species, or more likely a combination of both.
> * Scenario 3 shows the effect of preferring AquaMaps data over IUCN, while eliminating the presence threshold.  Just as a presence threshold of zero in scenario 1 drives a decrease in average score relative to the baseline, the zero threshold in scenario 3 drives a decrease in scores relative to scenario 2.  The large decrease seems to indicate that within the set of paired-map species, a zero threshold greatly increases the spatial representation of high-risk species relative to low-risk species.

>>>>

<!--

From PNAS http://www.pnas.org/site/authors/preparation.xhtml:

Only TIFF, EPS, and high-resolution PDF for Mac or PC are allowed for figures that will appear in the main text. Authors may submit PRC or U3D files for 3D images; these must be accompanied by 2D representations in TIFF, EPS, or high-resolution PDF format. (See SI below for supplementary material.) Color images must be in RGB (red, green, blue) mode. Include the font files for any text. Images must be final size, preferably one column width (8.7 cm). Figures wider than one column should be sized to 11.4 cm or 17.8 cm wide. Numbers, letters, and symbols should be no smaller than 6 points (2 mm) and no larger than 12 points (6 mm) after reduction and must be consistent. Composite figures must be preassembled. Figures must be submitted as separate files, not embedded in manuscript text. See the Digital Art Guidelines.

See also: http://www.pnas.org/site/misc/digitalart.pdf

Label like so: 
Fig. 2.
Diffraction patterns at three different angles and time-dependent structural changes of JDF-L1. (A–C) Diffraction images observed at different zone axes before time 0; n is the electron direction. (D) Femtosecond intensity changes of (200) and (201) spots with time. Two different types of dynamics were observed: a decay with a femtosecond time constant (see main text) and another without a decay.

@@

### Fig. 1

@@

>  __Fig. 1.__ (A) Number and proportion of species, listed by taxa, included in each dataset: IUCN, AquaMaps, or both.  AquaMaps encompasses a broader range of taxa than IUCN, while IUCN focuses on comprehensively assessing select taxonomic groups, typically at the level of order or family. Overlapping species are dominated by bony fishes (`r n_bonyfish` species, primarily tropical taxa) and corals (`r n_corals` species).  (B, C) Global marine species richness according to (B) AquaMaps dataset and (C) IUCN dataset.  The frequency plot to the right of each map shows relative species count per cell at each latitude; while both datasets peak in tropical latitudes near the equator, the frequency for IUCN maps drops quickly beyond 30°N and 30°S, while the frequency for AquaMaps remains robust well into temperate latitudes.  The frequency plot above each map shows relative species count at each longitude, showing a slight bias in the IUCN dataset away from the Atlantic and eastern Pacific compared to AquaMaps.

@@

![](figures/AreaRatiovsPercOverlap.png)
![](figures/barchart_spp_gp_quads.png)

@@

> __Fig. 2.__ (A) DESCRIPTION, then explanation... The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the extent of described ranges
(n = `r q1_count`; `r round(q1_count/nrow(spp_map_pairs) * 100, 1)` %).
> The upper left quadrant (quadrant 2) comprises species whose maps agree well in distribution, but disagree in extent
(n = `r q2_count`; `r round(q2_count/nrow(spp_map_pairs) * 100, 1)` %).
> The lower right quadrant (quadrant 3) includes species for which the paired maps generally agree in range extent, but disagree on where those ranges occur
(n = `r q3_count`; `r round(q3_count/nrow(spp_map_pairs) * 100, 1)` %).
> The lower left quadrant (quadrant 4) indicates species for which the map pairs agree poorly in both area and distribution
(n = `r q4_count`; `r round(q4_count/nrow(spp_map_pairs) * 100, 1)` %).  
(B) Spatial alignment of paired-map species by taxonomic group. _is it distracting to use order names for some groups and not others?  Some of the groups had pretty long names ("acanthuridae" is way shorter than "surgeonfish, tangs, and unicornfish")_

@@

> __Fig. 3.__ Change in status score for the Species Subgoal within the global Ocean Health Index under three different scenarios.

@@

> * Scenario 1 shows the effect of reducing the presence threshold for AquaMaps presence.  Reducing the threshold increases the apparent range of a species; the slight decrease in average score suggests increased spatial representation of threatened species.
> * Scenario 2 shows the effect of prioritizing AquaMaps data over IUCN, while maintaining the same presence threshold.  This will have different effects depending on the species; in general, AquaMaps ranges are smaller than IUCN ranges, so most overlapping species will see a decrease in represented range.  The slight bump in mean score may indicate a small increase in spatial representation of low-risk species, a small decrease in spatial representation of high-risk species, or more likely a combination of both.
> * Scenario 3 shows the effect of prioritizing AquaMaps data over IUCN, while eliminating the presence threshold.  Just as a presence threshold of zero in scenario 1 drives a decrease in average score relative to the baseline, the zero threshold in scenario 3 drives a decrease in scores relative to scenario 2.  The large decrease indicates that within the set of paired-map species, a zero threshold greatly increases the spatial representation of high-risk species relative to low-risk species.

### 2016-01-23 ms_6_supplemental.Rmd
updating the manuscript...

>>>>

from PNAS http://www.pnas.org/site/authors/preparation.xhtml:

The main text of the paper must stand on its own without the SI. Refer to SI in the manuscript at an appropriate point in the text. Number supporting figures and tables starting with S1, S2, etc. Authors are limited to no more than 10 SI files, not including movie files.

Authors who place detailed materials and methods in SI must provide sufficient detail in the main text methods to enable a reader to follow the logic of the procedures and results and also must reference the online methods. If a paper is fundamentally a study of a new method or technique, then the methods must be described completely in the main text.

Because PNAS edits SI and composes it into a single PDF, authors must provide the following file formats only:

Text: Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a PDF with the same file name for visual reference).
Figures: Provide a brief legend for each supporting figure after the supporting text. Provide figure images in TIFF, EPS, high-resolution PDF, JPEG, or GIF format; figures may not be embedded in manuscript text. When saving TIFF files, use only LZW compression; do not use JPEG compression. Do not save figure numbers, legends, or author names as part of the image. Composite figures must be preassembled.
3D figures: Supply a composable U3D or PRC file so that it may be edited and composed. Authors may submit a PDF file but please note it will be published in raw format and will not be edited or composed.
Tables: Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a PDF with the same file name for visual reference); include only one table per file. Do not use tabs or spaces to separate columns in Word tables.
Datasets: Supply Excel (.xls), RTF, or PDF files. This file type will be published in raw format and will not be edited or composed.
Movies: Supply Audio Video Interleave (avi), Quicktime (mov), Windows Media (wmv), animated GIF (gif), or MPEG files and submit a brief legend for each movie in a Word or RTF file. All movies should be submitted at the desired reproduction size and length. Movies should be no more than 10 MB in size.
Still images: Authors must provide a still image from each video file. Supply TIFF, EPS, high-resolution PDF, JPEG, or GIF files.
Appendices: PNAS prefers that authors submit individual source files to ensure readability. If this is not possible, supply a single PDF file that contains all of the SI associated with the paper. This file type will be published in raw format and will not be edited or composed.
